Analytic Occultation Light Curves in Reflected Light

### 2020-04-06 tex/ms.tex
more text

Similarly, the flux outside of an occultation is now given by
%
\begin{align}
    \label{eq:rTIA1RRy}
    f_0 & =
    \mathfrak{r}^\top(b)
    \BF{I}(b, \theta'', r_s)
    \BF{A_1}
    \BF{R}''(x_s, y_s)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:theta''}
    \theta'' = 0
\end{align}
%
is the angle of the terminator in the frame $\mathcal{F}''$, by
construction. The transformation
$\BF{R}'' = \BF{R}''(x_s, y_s)$ (Equation~\ref{eq:R''}) rotates
the body so the semi-major axis of the terminator is aligned
with the $x''$-axis; as will become clear in \S\ref{sec:solution-no-occ} below,
this greatly simplifies the integration step.

@@

Cases 6--10 (see Figure~\ref{fig:cases}) correspond to configurations in which
the limb of the occultor intersects with the terminator at either one point
(case 6) or two points (cases 7--10). Because of these intersections, we
cannot simply re-weight the emitted light solution, as the integration
boundaries are now different. In general, we may compute the flux by
integrating the components of the Green's basis $\bg$
over the region $S$ bounded by three curves, which we denote
$P$, $T$, and $Q$. These are shown in orange
in Figure~\ref{fig:cases} and presented in more detail in
Figure~\ref{fig:geometry}.
Curve $P$ is a segment of the limb of
the occultor, parametrized by the angle $\phi \in [\phi_1, \phi_2]$;
curve $T$ is a segment of the terminator,
parametrized by the angle $\xi \in [\xi_1, \xi_2]$;
and curve $Q$ is a segment of the limb of the occulted body,
parametrized by the angle $\lambda \in [\lambda_1, \lambda_2]$.
The endpoints $\phi_1, \phi_2, \xi_1, \xi_2, \lambda_1,$ and
$\lambda_2$ are functions of the solutions to the quartic from
\S\ref{sec:which-case} and will be presented below.

As in \citet{Luger2019}, the approach to computing the surface integrals
of each of the terms in $\bg$ over $S$
is to use Green's theorem to transform them into line integrals along the
curves $P$, $T$, and $Q$. Let $\mathfrak{p}^\top$,
$\mathfrak{t}^\top$, and $\mathfrak{q}^\top$ denote the vectors
corresponding to the line integrals
along the curves $P$, $T$, and $Q$, respectively, of the anti-exterior
derivatives of each of the terms in
$\bg$ (see Appendix~\ref{sec:sT}).
The integration is performed in the frame $\mathcal{F}'$,
in a counter-clockwise direction relative to the center
of the integration region $S$.
Then the integral of $\bg$ over $S$ is%
\footnote{%
    The components of the vectors $\mathfrak{p}^\top$ and
    $\mathfrak{q}^\top$ are analogous to the ``primitive integrals''
    $\mathcal{P}$ and $\mathcal{Q}$
    defined in Equations~(30)--(32) in \citet{Luger2019}, although the
    integration limits of both (and the sense of integration of $\mathcal{P}$)
    are different.
}
%
\begin{align}
    \label{eq:sT}
    \sT & =
    \mathfrak{p}^\top + \mathfrak{t}^\top + \mathfrak{q}^\top
    \quad,
\end{align}
%
where the dependence on $b, \theta', b_o,$ and $r_o$ is implicit.
%
We defer the solution to Equation~(\ref{eq:sT}) to \S\ref{sec:sT} below,
as it is quite lengthy. Given $\sT$, the flux $f_S$ over the integration
region $S$ is computed from Equation~(\ref{eq:sTA2IA1RRy}):
%
\begin{align}
    \label{eq:fS}
    f_S & =
    \mathfrak{s}^\top(b, \theta', b_o, r_o)
    \BF{A_2}
    \BF{I}(b, \theta', r_s)
    \BF{A_1}
    \BF{R}'(x_o, y_o)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad.
\end{align}
%
Note again that this is not necessarily the \emph{observed} flux, which we must
compute on a case-by-case basis below.

Case 6 corresponds to configurations in which the limb of the occultor
intersects the terminator at a single point. The integration region
(see Figures~\ref{fig:cases} and \ref{fig:geometry}) is the occulted
portion of the dayside, which is bounded by all three curves $P$, $Q$, and $T$.
The total flux may
be computed by subtracting the occulted flux $f_S$ from the total dayside
flux $f_0$:
%
\begin{proof}{}
    \label{eq:f6}
    f_6 = f_0 - f_S
    \quad.
\end{proof}
%
Cases 7--10 involve two points of intersection between the occultor limb and
the terminator. The region we integrate over is bounded only by the terminator
and the limb of the occultor, so when computing $f_S$ below we set
$\mathfrak{q}^\top = \boldsymbol{0}$.
%
Cases 7 and 8 correspond to occultors that block some of the
nightside and some of the dayside, but \emph{neither} of the extrema of the
terminator ellipse. In case 7 a lens-shaped region is formed by the
intersection of the occultor limb and the terminator on the
\emph{nightside}, while in case 8 this region is formed on the \emph{dayside}.
%
In case 7, we begin by computing the
flux over the unocculted region, $f_I$, which includes the spurious
nightside contribution. We then remove this contribution by noting that it
is equal to the total nightside contribution, $\hat{f}_0$, minus the
occulted nightside flux, $f_S$:
%
\begin{proof}{}
    \label{eq:f7}
    f_7 = f_I - (\hat{f}_0 - f_S)
    \quad.
\end{proof}
%
Case 8, on the other hand, is topologically equivalent to case 6, since
the integration region consists of occulted dayside:
%
\begin{proof}{}
    \label{eq:f8}
    f_8 = f_0 - f_S
    \quad.
\end{proof}
%
Cases 9 and 10 correspond to occultors that also block some nightside and some
dayside, along with \emph{both} of the extrema of the ellipse; these are
therefore exclusively for large occultors ($r_o > 1$). Case 9
involves occultations in which only a small lens-shaped region of the
nightside is visible. The total flux is the visible dayside plus unphysical
nightside contribution, $f_I$, minus the nightside contribution, which we
compute from Equation~(\ref{eq:fS}):
%
\begin{proof}{}
    \label{eq:f9}
    f_9 = f_I - f_S
    \quad.
\end{proof}
%
Conversely, case 10 involves occultations in which only a small lens-shaped
region of the dayside is visible. In this case, we may compute the observed
flux from Equation~(\ref{eq:fS}) directly:
%
\begin{proof}{}
    \label{eq:f10}
    f_{10} = f_S
    \quad.
\end{proof}
%

@@

Cases 11--14 correspond to (rare) configurations involving three
or four roots to Equation~\ref{eq:quartic} and are
illustrated in Figure~\ref{fig:pathological}. All four involve integration over
two disjoint regions (see the figure).
%
Cases 11 and 12 involve three points of intersection between the terminator
and the occultor limb. In case 11, the regions of integration $S_1$ and
$S_2$ are the
occulted portion of the dayside, so the solution is similar to
that of cases 6 and 8:
%
\begin{proof}{}
    \label{eq:f11}
    f_{11} = f_0 - (f_{S_1} + f_{S_2})
    \quad,
\end{proof}
%
where $f_{S_1}$ and $f_{S_2}$ are computed from Equation~(\ref{eq:fS}) for
each of the integration regions. Conversely, in case 12 the two regions
are the occulted portion of the nightside, so the solution is similar to that
of case 7:
%
\begin{proof}{}
    \label{eq:f12}
    f_{12} = f_I - \big(\hat{f}_0 - (f_{S_1} + f_{S_2})\big)
    \quad,
\end{proof}
%
Finally, cases 13 and 14 involve four points of intersection between the
terminator and the occultor limb. The regions of integration in case 13 are
the visible portion of the nightside, so this case is equivalent to case 9:
%
\begin{proof}{}
    \label{eq:f13}
    f_{13} = f_I - (f_{S_1} + f_{S_2})
    \quad.
\end{proof}
%
Conversely, the regions of integration in case 14 are the
visible portion of the dayside, so this case is equivalent to case 10:
%
\begin{proof}{}
    \label{eq:f14}
    f_{14} = (f_{S_1} + f_{S_2})
    \quad.
\end{proof}
%

\subsection{Computing the integrals $\sT$}
\label{sec:sT}
\xxx{This is going to be be painful.}

>>>>

Similarly, the flux when there is no occultation is now given by
%
\begin{align}
    \label{eq:rTIA1RRy}
    f_0 & =
    \mathfrak{r}^\top(b)
    \BF{I}(b, \theta'', r_s)
    \BF{A_1}
    \BF{R}''(x_s, y_s)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:theta''}
    \theta'' = 0
\end{align}
%
is the angle of the terminator in the frame $\mathcal{F}''$, by
construction. The transformation
$\BF{R}'' = \BF{R}''(x_s, y_s)$ (Equation~\ref{eq:R''}) rotates
the body so the semi-major axis of the terminator is aligned
with the $x''$-axis; as will become clear in \S\ref{sec:solution-no-occ} below,
this greatly simplifies the integration step.

@@

Cases 6--10 (see Figure~\ref{fig:cases}) correspond to configurations in which
the limb of the occultor intersects with the terminator at either one point
(case 6) or two points (cases 7--10). Because of these intersections, we
cannot simply re-weight the emitted light solution, as the integration
boundaries are now different. In general, we may compute the flux by
integrating the components of the Green's basis $\bg$
over the region $S$ bounded by three curves, which we denote
$P$, $T$, and $Q$. These are shown in orange
in Figure~\ref{fig:cases} and presented in more detail in
Figure~\ref{fig:geometry}.
Curve $P$ is a segment of the limb of
the occultor, parametrized by the angle $\phi \in [\phi_1, \phi_2]$;
curve $T$ is a segment of the terminator,
parametrized by the angle $\xi \in [\xi_1, \xi_2]$;
and curve $Q$ is a segment of the limb of the occulted body,
parametrized by the angle $\lambda \in [\lambda_1, \lambda_2]$.
The endpoints $\phi_1, \phi_2, \xi_1, \xi_2, \lambda_1,$ and
$\lambda_2$ are functions of the solutions to the quartic from
\S\ref{sec:which-case} and will be presented in \S\ref{sec:sT}.

Let $\sT$ be the integral of $\bg^\top$ over $S$:
%
\begin{align}
    \label{eq:sTint}
    \sT(b, \theta', b_o, r_o) & =
    \iint\limits_{S(b, \theta', b_o, r_o)}
    \bg^\top(x', y')
    \ \dd x' \ \dd y'
    \quad,
\end{align}
%
We defer the solution to Equation~(\ref{eq:sTint}) to \S\ref{sec:sT} below,
as it is quite lengthy. Given $\sT$, the flux $f_S$ over the integration
region $S$ is computed from Equation~(\ref{eq:sTA2IA1RRy}):
%
\begin{align}
    \label{eq:fS}
    f_S & =
    \mathfrak{s}^\top(b, \theta', b_o, r_o)
    \BF{A_2}
    \BF{I}(b, \theta', r_s)
    \BF{A_1}
    \BF{R}'(x_o, y_o)
    \BF{R}(i, \lambda, \vartheta)
    \BF{y}
    \quad.
\end{align}
%
Note again that this is not necessarily the \emph{observed} flux, which we must
compute on a case-by-case basis below.

Case 6 corresponds to configurations in which the limb of the occultor
intersects the terminator at a single point. The integration region
(see Figures~\ref{fig:cases} and \ref{fig:geometry}) is the occulted
portion of the dayside, which is bounded by all three curves $P$, $Q$, and $T$.
The total flux may
be computed by subtracting the occulted flux $f_S$ from the total dayside
flux $f_0$:
%
\begin{proof}{}
    \label{eq:f6}
    f_6 = f_0 - f_S
    \quad.
\end{proof}
%
Cases 7--10 involve two points of intersection between the occultor limb and
the terminator.
%
Cases 7 and 8 correspond to occultors that block some of the
nightside and some of the dayside, but \emph{neither} of the extrema of the
terminator ellipse. In case 7 a lens-shaped region is formed by the
intersection of the occultor limb and the terminator on the
\emph{nightside}, while in case 8 this region is formed on the \emph{dayside}.
%
In case 7, we begin by computing the
flux over the unocculted region, $f_I$, which includes the spurious
nightside contribution. We then remove this contribution by noting that it
is equal to the total nightside contribution, $\hat{f}_0$, minus the
occulted nightside flux, $f_S$:
%
\begin{proof}{}
    \label{eq:f7}
    f_7 = f_I - (\hat{f}_0 - f_S)
    \quad.
\end{proof}
%
Case 8, on the other hand, is topologically equivalent to case 6, since
the integration region consists of occulted dayside:
%
\begin{proof}{}
    \label{eq:f8}
    f_8 = f_0 - f_S
    \quad.
\end{proof}
%
Cases 9 and 10 correspond to occultors that also block some nightside and some
dayside, along with \emph{both} of the extrema of the ellipse; these are
therefore exclusively for large occultors ($r_o > 1$). Case 9
involves occultations in which only a small lens-shaped region of the
nightside is visible. The total flux is the visible dayside plus unphysical
nightside contribution, $f_I$, minus the nightside contribution, which we
compute from Equation~(\ref{eq:fS}):
%
\begin{proof}{}
    \label{eq:f9}
    f_9 = f_I - f_S
    \quad.
\end{proof}
%
Conversely, case 10 involves occultations in which only a small lens-shaped
region of the dayside is visible. In this case, we may compute the observed
flux from Equation~(\ref{eq:fS}) directly:
%
\begin{proof}{}
    \label{eq:f10}
    f_{10} = f_S
    \quad.
\end{proof}
%

@@

Cases 11--14 correspond to (rare) configurations involving three
or four roots to Equation~\ref{eq:quartic} and are
illustrated in Figure~\ref{fig:pathological}. All four involve integration over
two disjoint regions (see the figure).
%
Cases 11 and 12 involve three points of intersection between the terminator
and the occultor limb. In case 11, the regions of integration $S_1$ and
$S_2$ are the
occulted portion of the dayside, so the solution is similar to
that of cases 6 and 8:
%
\begin{proof}{}
    \label{eq:f11}
    f_{11} = f_0 - (f_{S_1} + f_{S_2})
    \quad,
\end{proof}
%
where $f_{S_1}$ and $f_{S_2}$ are computed from Equation~(\ref{eq:fS}) for
each of the integration regions. Conversely, in case 12 the two regions
are the occulted portion of the nightside, so the solution is similar to that
of case 7:
%
\begin{proof}{}
    \label{eq:f12}
    f_{12} = f_I - \big(\hat{f}_0 - (f_{S_1} + f_{S_2})\big)
    \quad,
\end{proof}
%
Finally, cases 13 and 14 involve four points of intersection between the
terminator and the occultor limb. The regions of integration in case 13 are
the visible portion of the nightside, so this case is equivalent to case 9:
%
\begin{proof}{}
    \label{eq:f13}
    f_{13} = f_I - (f_{S_1} + f_{S_2})
    \quad.
\end{proof}
%
Conversely, the regions of integration in case 14 are the
visible portion of the dayside, so this case is equivalent to case 10:
%
\begin{proof}{}
    \label{eq:f14}
    f_{14} = f_{S_1} + f_{S_2}
    \quad.
\end{proof}
%

\subsection{Computing the integrals $\sT$}
\label{sec:sT}
%
In the previous sections, we discussed how to identify the case
corresponding to a specific configuration of the occultor and the
illumination source. We showed how in some cases
(1--5; \S\ref{sec:cases-easy}) the total flux may be
computed by exploiting the classical \starry integrals (Equation~\ref{eq:fI}).
In all other cases (6--14; \S\ref{sec:cases-hard} and
\S\ref{sec:cases-pathological}), however, the flux computation involves
evaluation of
Equation~(\ref{eq:fS}), where the solution vector $\mathfrak{s}^\top$
is the vector of integrals of each of the terms in the Green's basis $\bg$
over a region $S$ of the
projected disk of the occultor (Equation~\ref{eq:sTint}).
%
As in \citet{Luger2019}, the approach to computing $\sT$
is to use Green's theorem to transform the surface integrals into line
integrals along the curves $P$, $T$, and $Q$
(see Figure~\ref{fig:geometry}). Specifically, we write%
\footnote{%
    In this section, we deliberately drop the prime symbols from $x'$ and
    $y'$ for clarity. We also drop the dependence of $\sT$ and the primitive
    integrals on the geometrical parameters $b, \theta', b_o, r_o$.
}
%
\begin{align}
    \label{eq:greens}
    \sT
     & =
    \iint\limits_{S}
    \bg^\top(x, y)
    \ \dd x \ \dd y
    \nonumber \\[0.5em]
     & =
    \oint \bvec{G} (x, y) \cdot
    \dd \bvec{r} (x, y)
    \quad,
\end{align}
%
where $\bvec{G} (x, y)$
is an array of two-dimensional Cartesian vectors chosen such that its
exterior derivative is $\bg$,
%
\begin{align}
    \label{eq:DGg}
    \frac{\dd \bvec{G}_{y}(x, y)}{\dd x}
    - \frac{\dd \bvec{G}_{x}(x, y)}{\dd y} = \bg(x, y)
    \quad,
\end{align}
%
and
%
\begin{align}
    \dd \BF{r} (x, y) & =
    \left(\frac{\dd x}{\dd \varphi}\right) \dd \varphi \, \xhat +
    \left(\frac{\dd y}{\dd \varphi}\right) \dd \varphi \, \yhat
    \quad,
\end{align}
%
where $\varphi$ is the parametrized angle along the integration path.
%
\citet{Luger2019} showed that one possible solution to
Equation~(\ref{eq:DGg}) consists of the vector whose $n^\mathrm{th}$
component is given by
%
\begin{align}
    \label{eq:G}
    \BF{G}_n (x, y) & =
    \begin{dcases}
        %
        x^{\frac{\mu + 2}{2}}
        y^{\frac{\nu}{2}}
        \,\yhat
         & \qquad \mu, \nu \, \mathrm{even}
        \\[1em]
        %
        \frac{1-z(x, y)^3}{3(1-z(x, y)^2)}\bigg(-y \, \xhat + x \, \yhat\bigg)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        x^{l-2}
        z(x, y)^3
        \,\xhat
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{even}
        \\[1em]
        %
        x^{l-3}
        y
        z(x, y)^3
        \,\xhat
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{odd}
        \\[1em]
        %
        x^{\frac{\mu-3}{2}}
        y^{\frac{\nu-1}{2}}
        z(x, y)^3
        \,\yhat
         & \qquad \mathrm{otherwise,}
    \end{dcases}
\end{align}
%
where the indices $l, m, \mu, \nu$ are given by Equation~(\ref{eq:l-m-mu-nu}).

We showed in the previous sections that there are at most three curves
$P$, $T$, and $Q$ bounding a given closed surface of integration
(see Figure~\ref{fig:geometry}). We may therefore express
Equation~(\ref{eq:greens}) as
%
\begin{align}
    \label{eq:sT}
    \sT & =
    \mathfrak{p}^\top + \mathfrak{t}^\top + \mathfrak{q}^\top
    \quad,
\end{align}
%
where we define the primitive integrals
%
\begin{align}
    \label{eq:pT}
    \mathfrak{p}^\top
     & =
    \int\limits_{\phi_1}^{\phi_2}
    \mathbf{G}(x_p, y_p)
    \cdot \dd \mathbf{r}(x_p, y_p)
    %
    \\
    %
    \label{eq:tT}
    \mathfrak{t}^\top
     & =
    \int\limits_{\xi_1}^{\xi_2}
    \mathbf{G}(x_t, y_t)
    \cdot \dd \mathbf{r}(x_t, y_t)
    %
    \\
    %
    \label{eq:qT}
    \mathfrak{q}^\top
     & =
    \int\limits_{\lambda_1}^{\lambda_2}
    \mathbf{G}(x_q, y_q)
    \cdot \dd \mathbf{r}(x_q, y_q)
\end{align}
%
to be the line integrals of $\mathbf{G}$ along each of the curves
$P$, $T$, and $Q$, respectively, where the coordinates are parametrized
in terms of $\varphi$ as follows:
%
\\[1em]
%
\begin{minipage}{0.31\linewidth}
    \begin{align}
        x_p & = r_o \cos\varphi
        \nonumber                     \\
        y_p & = b_o + r_o \sin\varphi
        \nonumber
    \end{align}
\end{minipage}
%
\begin{minipage}{0.35\linewidth}
    \begin{align}
        x_t & = \cos\theta' \cos\varphi - b \sin\theta' \sin\varphi
        \nonumber                                                   \\
        y_t & = \sin\theta' \cos\varphi + b \cos\theta' \sin\varphi
        \nonumber
    \end{align}
\end{minipage}
%
\begin{minipage}{0.31\linewidth}
    \begin{align}
        \label{eq:xy_pqt}
        x_q & =\cos\varphi
        \nonumber           \\
        y_q & = \sin\varphi
        \quad.
    \end{align}
\end{minipage}
%
\\[1em]

\xxx{We perform the integration in a frame $\mathcal{F}'$
    in which the occultor is aligned with the $+y'$-axis,
    in a counter-clockwise direction relative to the center
    of the integration region $S$.}
